---
title: "TakeHome_Ex03"
author: "Xu Jiajie"
---

# 1.0 Task

In this take-home exercise, I need to uncover the salient patterns of the resale prices of public housing. We need to focus on 3-Room, 4-Room, and 5-room types. I would focus on different factors' influence on resale price and after that, I would focus on analyzing the resale price from one type of house.

# 2.0 The data

***Resale flat princes based on registration date from Jan-2017 onwards*** should be used to prepare the analytical visualization. It is available at [Data.gov.sg](https://data.gov.sg/).

# 3.0 Import the dataset and prepare the data

3.1 Installing and loading the required libraries

```{r}
pacman::p_load(tidyverse, patchwork, ggthemes, hrbrthemes, ggrepel, ggiraph, ggstatsplot, PMCMRplus, performance, parameters, see, plotly, factoextra,ggplot2,forcats,dendextend,heatmaply,pheatmap,rmarkdown)
```

3.2 Importing the dataset and preparing the data

```{r}
Flat_resale <- read_csv("data/2017-2022.csv")
Flat_resale
```

3.3 Because we need to focus on 3-room, 4-room, and 5-room house type so I fiter out the data that we need.

```{r}
Flat_resale1 <- filter(Flat_resale,flat_type=="3 ROOM"|flat_type=="4 ROOM"|flat_type=="5 ROOM")
```

```{r}
Flat_resale1
```

3.4 The task require us to analysis 2022's data so I filter 2022's data. The following steps can filter out the data that we need.

```{r}
Flat_resale2 <- Flat_resale1 %>% 
  separate(month, c("year", "months"), "-")
Flat_resale2
```

```{r}
Flat_resale3 <- filter(Flat_resale2,year=="2022")
Flat_resale3
```

## 4. Creating the Visualisation

4.1 I want to compare differnt house type's resale_price's median. But I need to decrease the influence of house size. So, I create one more column to store the average price/m² and named it AvgPrice.

```{r}
Flat_resale4 <- mutate(Flat_resale3,AvgPrice=resale_price/floor_area_sqm)
Flat_resale4
```

4.2 Two-sample mean test:

In the code chunk below, [**ggbetweenstats()**](https://indrajeetpatil.github.io/ggstatsplot/reference/ggbetweenstats.html) is used to build a visual for two-sample mean test of AvgPrice by house types.

```{r}
ggbetweenstats(
  data = Flat_resale4,
  x =flat_type, 
  y =AvgPrice,
  type = "np",
  messages = FALSE
)
```

From above chart, it is easy to find that 4-room's AvgPrice is higher than others. It means in the same house size, 4-room house would more expensive than others and the 3-room house is more expensive than 5-room house.

4.3 checking normality assumption between AvgPrice and storey_range.

```{r}
model1 <- lm(AvgPrice ~storey_range , data = Flat_resale4)
model1
```

```{r}
check_n <- check_normality(model1)
```

```{r}
plot(check_n)
```

4.4 Oneway ANOVA Test: Do a ANOVA Test to verify that is there any difference in AvgPrice based on different storeys.

H0: There is irrelevant between AvgPrice and storey_range.

H1: There is relevant between AvgPrice and storey_range.

```{r}
ggbetweenstats(
  data = Flat_resale4,
  x = storey_range, 
  y = AvgPrice,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

From above result, we can see that P\<0.05. So we can reject H0. So, AvgPrice depends on the floor. The height of the floor affects the average price.

4.5 Exploring the relationship between resale_price and storey_range.

In order to explore more about the resale_price and storey_range.

```{r}
exam1 <- Flat_resale4 %>% 
  mutate(AvgPrice_bin = 
           cut(AvgPrice, 
               breaks = c(0,2000,4000,6000,8000))
)
```

```{r}
ggbarstats(exam1, 
           x = AvgPrice_bin, 
           y = storey_range)
```

From above chart, we can see that before storey 37, the higher storey , the green color occupied more. Which means the higher storey, the higher AvgPrice.

## 5. Exploring one of house type: 4-Room type.

Because this house type is very popular in SG so we want to explore more about this type.

5.1 Data prepare: Filter out the 4-Room type data.

```{r}
Flat_resale5 <- filter(Flat_resale4,flat_type=="4 ROOM")
Flat_resale5
```

5.2 The median of resale_price in different towns.

In order to explore the price in different towns, we need to group by every town and calculate the median of resale_price:

```{r}
Flat_resale6 <- Flat_resale5 %>%
                group_by(town) %>%
                summarize(median_price = median(resale_price))
Flat_resale6
```

```{r}
ggplot(data=Flat_resale6,aes(x=reorder(town,median_price),y=median_price))+
      geom_bar(stat = "identity")
```

From above data and chart, we can find that some towns have a really high median of resale_price, although all of them are 4-room house. We can explore step further.

5.3 Visualizing the uncertainty of point estimates

```{r}
my_sum <- Flat_resale5 %>%
  group_by(town) %>%
  summarise(
    n=n(),
    mean=mean(resale_price),
    sd=sd(resale_price)
    ) %>%
  mutate(se=sd/sqrt(n-1))
```

```{r}
knitr::kable(head(my_sum), format = 'html')
```

```{r}
ggplot(my_sum) +
  geom_errorbar(
    aes(x=town, 
        ymin=mean-se, 
        ymax=mean+se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=town, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle("Standard error of mean 
          resale_price score by rac")
```

**Model Diagnostic: checking normality assumption**

In the code chunk, [*check_normality()*](https://easystats.github.io/performance/reference/check_normality.html) of [**performance**](https://easystats.github.io/performance/) package.

```{r}

area_price <- lm(floor_area_sqm ~resale_price , data = Flat_resale5)
```

```{r}
check_n <- check_normality(area_price)
```

```{r}
plot(check_n)
```

```{r}
ggscatterstats(
  data = Flat_resale5,
  x =floor_area_sqm ,
  y = resale_price,
  marginal = FALSE,
  )
```
